---
title: Protein analysis
author: "AS Huffmyer"
date: '2022'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for total protein for the E5 *A. pulchra* metabolism 2022 project. Scripts based on E5 timeseries analysis scripts. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("broom")) install.packages("broom")
if (!require("ggpubr")) install.packages("ggpubr")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")

# load packages
library(tidyverse)
library(broom)
library(ggpubr)
library(lme4)
library(lmerTest)
```

# Host protein analysis  

## Import data for host protein
```{r}
prot_host<-read_csv("data/baseline_sampling/protein/20221013_host_plate1.csv")
prot_holobiont<-read_csv("data/baseline_sampling/protein/20221013_holobiont_plate2.csv")
platemap<-read_csv("data/baseline_sampling/protein/20221013_platemap.csv")

# Merge platemap and data for host data 
df_host <- left_join(prot_host, platemap, by="well")
```

## Plot standard curve
```{r}
# Create standard curve following kit instructions
standards_host <- tribble(
  ~std, ~BSA_ug.mL,
  "A",        2000,
  "B",        1500,
  "C",        1000,
  "D",         750,
  "E",         500,
  "F",         250,
  "G",         125,
  "H",          25,
  "I",           0
)


std_curve_host <- df_host %>%
  filter(grepl("Standard", fragment)) %>%
  select(well, fragment, abs562 = `562:562`) %>%
  rename(std = fragment) %>%
  mutate(std = str_sub(std, 9, 9)) %>%
  #group_by(std) %>%
  #summarise(abs562 = mean(abs562)) %>%                       # calculate mean of standard duplicates
  #mutate(abs562.adj = abs562 - abs562[std == "I"]) %>%       # subtract blank absorbace value from all
  left_join(standards_host)

## Fit linear model for standard curve
mod_host <- lm(BSA_ug.mL ~ abs562, data = std_curve_host)
coef(mod_host)

fitted_host <- mod_host %>% broom::augment()
```

Plot the standard curve with linear statistics.  
```{r}
standard.plot.host <- ggplot(data = std_curve_host, aes(x=BSA_ug.mL, y=abs562))+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));standard.plot.host
```

## Calculate protein concentrations

Calculate protein and project on the standard curve.  
```{r}
# Calculate protein concentration for all samples using standard curve
prot_host <- df_host %>%
  filter(!grepl("Standard", fragment)) %>%                     # Get just samples (not standards)
  select(well, fragment, abs562 = `562:562`) %>%        # Select only needed columns
  filter(!is.na(fragment)) %>%                                 # Filter out empty wells
  #filter(fragment != "BK") %>%                                 # Filter out blank wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod_host, newdata = data.frame(abs562 = .))))    # Use standard curve to convert absorbance to protein

# Plot standard curve
std_curve_plot_host <- std_curve_host %>%
  ggplot(aes(x = abs562, y = BSA_ug.mL)) +
  geom_point(color = "red", size = 3) 

std_curve_plot_host + 
  geom_point(data = prot_host, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve")
```

## Normalize to surface area
```{r}
# Surface area data
sa <- read.csv("output/baseline/surface_area_calc.csv")

# Tissue homogenate volume data
homog_vols <- read_csv("data/baseline_sampling/homog_vols.csv") %>% select(1:2)

# Coral sample metadata
metadata <- read_csv("data/fragments.csv") %>% select(1:4) %>% filter(treatment=="Baseline")

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols) %>%
  full_join(sa)

# Join host prot data with metadata
prot_host <- left_join(prot_host, metadata) %>%
  mutate(prot_ug_host = prot_ug.mL * homog.vol.mL,
         prot_ug.cm2_host = prot_ug_host / surface.area.cm2,
         prot_mg.cm2_host = prot_ug.cm2_host / 1000)%>%
  select(-abs562, -prot_ug.mL)
```

## Plot results by genotype
```{r}
# Plot all data points with mean ± se
prot_host %>%
  filter(!is.na(colony))%>%
  ggplot(aes(x = colony, y = prot_mg.cm2_host, color = colony)) +
  coord_cartesian(ylim = c(0, 1.0))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean
```

# Holobiont protein analysis  

## Import data for host protein
```{r}
# Merge platemap and data for host data 
df_holobiont <- left_join(prot_holobiont, platemap, by="well")
```

## Plot standard curve
```{r}
# Create standard curve following kit instructions
standards_holobiont <- tribble(
  ~std, ~BSA_ug.mL,
  "A",        2000,
  "B",        1500,
  "C",        1000,
  "D",         750,
  "E",         500,
  "F",         250,
  "G",         125,
  "H",          25,
  "I",           0
)


std_curve_holobiont <- df_holobiont %>%
  filter(grepl("Standard", fragment)) %>%
  select(well, fragment, abs562 = `562:562`) %>%
  rename(std = fragment) %>%
  mutate(std = str_sub(std, 9, 9)) %>%
  #group_by(std) %>%
  #summarise(abs562 = mean(abs562)) %>%                       # calculate mean of standard duplicates
  #mutate(abs562.adj = abs562 - abs562[std == "I"]) %>%       # subtract blank absorbace value from all
  left_join(standards_holobiont)%>%
  filter(abs562<3)                # remove outlier standard point

## Fit linear model for standard curve
mod_holobiont <- lm(BSA_ug.mL ~ abs562, data = std_curve_holobiont)
coef(mod_holobiont)

fitted_holobiont <- mod_holobiont %>% broom::augment()
```

Plot the standard curve with linear statistics.  
```{r}
standard.plot.holobiont <- ggplot(data = std_curve_holobiont, aes(x=BSA_ug.mL, y=abs562))+
  ylab("Absorbance (nm)")+ xlab("Protein (ug/mL)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 1.0, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.75, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));standard.plot.holobiont
```

## Calculate protein concentrations

Calculate protein and project on the standard curve.  
```{r}
# Calculate protein concentration for all samples using standard curve
prot_holobiont <- df_holobiont %>%
  filter(!grepl("Standard", fragment)) %>%                     # Get just samples (not standards)
  select(well, fragment, abs562 = `562:562`) %>%        # Select only needed columns
  filter(!is.na(fragment)) %>%                                 # Filter out empty wells
  #filter(fragment != "BK") %>%                                 # Filter out blank wells
  mutate(prot_ug.mL = map_dbl(abs562, ~ predict(mod_holobiont, newdata = data.frame(abs562 = .))))    # Use standard curve to convert absorbance to protein

# Plot standard curve
std_curve_plot_holobiont <- std_curve_holobiont %>%
  ggplot(aes(x = abs562, y = BSA_ug.mL)) +
  geom_point(color = "red", size = 3) 

std_curve_plot_holobiont + 
  geom_point(data = prot_holobiont, aes(x = abs562, y = prot_ug.mL), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve")
```

All samples are within the standard curve.  

## Normalize to surface area
```{r}
# Surface area data
sa <- read.csv("output/baseline/surface_area_calc.csv")

# Tissue homogenate volume data
homog_vols <- read_csv("data/baseline_sampling/homog_vols.csv") %>% select(1:2)

# Coral sample metadata
metadata <- read_csv("data/fragments.csv") %>% select(1:4) %>% filter(treatment=="Baseline")

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols) %>%
  full_join(sa)

# Join host prot data with metadata
prot_holobiont <- left_join(prot_holobiont, metadata) %>%
  mutate(prot_ug_holobiont = prot_ug.mL * homog.vol.mL,
         prot_ug.cm2_holobiont = prot_ug_holobiont / surface.area.cm2,
         prot_mg.cm2_holobiont = prot_ug.cm2_holobiont / 1000)%>%
  select(-abs562, -prot_ug.mL)
```

## Plot results by genotype
```{r}
# Plot all data points with mean ± se
prot_holobiont %>%
  filter(!is.na(colony))%>%
  ggplot(aes(x = colony, y = prot_mg.cm2_holobiont, color = colony)) +
  coord_cartesian(ylim = c(0, 1.0))+
  labs(x = "", y = "Total holobiont protein (mg/cm2)") +
  geom_jitter(width = 0.1) +                                            # Plot all points
  stat_summary(fun.data = mean_cl_normal, fun.args = list(mult = 1),    # Plot standard error
               geom = "errorbar", color = "black", width = 0.5) +
  stat_summary(fun.y = mean, geom = "point", color = "black")           # Plot mean
```

# Merge holobiont and host protein data  

```{r}
protein<-left_join(prot_host, prot_holobiont)

protein<-protein %>%
  mutate(tank=as.factor(tank))%>%
  filter(!is.na(colony))%>%
  select(fragment, colony, tank, prot_ug.cm2_holobiont, prot_ug.cm2_host)%>%
  rename(holobiont=prot_ug.cm2_holobiont, host=prot_ug.cm2_host)%>%
  gather(key="fraction", value="protein_ug.cm2", holobiont:host) 
```

## Plot host and holobiont data together  

Plot by genotype.  
```{r}
# Plot all data points with mean ± se
protein_plot<-protein %>%
  
  ggplot(aes(x = colony, y = protein_ug.cm2, color = fraction)) +
  #coord_cartesian(ylim = c(0, 1.0))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_boxplot()+
  ylim(200,1000)+
  geom_point(pch = 21, position = position_jitterdodge(0.1))+
  theme_classic(); protein_plot # Plot all points

ggsave(protein_plot, file="figures/baseline/protein.png", width=6, height=6)
```

Plot by tank  
```{r}
# Plot all data points with mean ± se
protein_plot_tank<-protein%>%
  
  ggplot(aes(x = tank, y = protein_ug.cm2, color = fraction)) +
  #coord_cartesian(ylim = c(0, 1.0))+
  labs(x = "", y = "Total protein (mg/cm2)") +
  geom_boxplot()+
  ylim(200,1000)+
  geom_point(pch = 21, position = position_jitterdodge(0.1))+
  theme_classic(); protein_plot_tank # Plot all points
```

## Statistical tests 

Analyze by differences in genotype using a mixed effect model with tank as a random effect.  

Analyze host.  
```{r}
model_host<-protein%>%
  filter(fraction=="host")%>%
  
  lmer(protein_ug.cm2~colony+(1|tank), data=.)

anova(model_host)
```

Analyze holobiont  
```{r}
model_holobiont<-protein%>%
  filter(fraction=="holobiont")%>%
  
  lmer(protein_ug.cm2~colony+(1|tank), data=.)

anova(model_holobiont)
```

Both host and holobiont protein are different by colony.  

# Write output file of all data  

```{r}
# Write protein data to output file
protein %>%
  filter(!is.na(colony)) %>%
  group_by(fragment) %>%
  mutate(timepoint="Baseline")%>%
  select(-)
  write_csv(., file = "output/baseline/protein_calc.csv")
```


