---
title: Surface area - wax dipping method
author: "AS Huffmyer"
date: '2022'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

This script analyzes surface area calculated by the wax dipping methods. Scripts based on Danielle Becker-Polinski's scripts.  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")

# load packages
library(tidyverse)
```

# Calculate standards 

Calculate standard curve from wooden dowels.  
```{r}
#load wax data
wax.data <- read.csv("data/baseline_sampling/surface_area.csv", header=TRUE)

wax.data$delta.mass.g <- wax.data$wax.weight.g-wax.data$dry.weight.g
stnds <- subset(wax.data, sample=="Standard")

#calculate the surface area of the spherical standards from the diameter
stnds$rad <- stnds$diameter/2
stnds$surface.area.cm2 <- 4*pi*(stnds$rad)^2

# calculate the curve coefficients for slope and intercept to apply as the standard
stnd.curve <- lm(surface.area.cm2~delta.mass.g, data=stnds)
stnd.plot<-plot(surface.area.cm2~delta.mass.g, data=stnds);stnd.plot
stnd.curve$coefficients
summary(stnd.curve)$r.squared #r-squared = 0.979 20221014
```

Plot the curve.  
```{r}
standard.plot <- ggplot(data = stnds, aes(x=delta.mass.g, y=surface.area.cm2))+
  ylab("Surface area (cm2)")+ xlab("Delta mass (g)") + 
  geom_point()+
  geom_smooth(method = "lm") +
  stat_regline_equation(label.y = 15.0, label.x = 3, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 5.0, label.x = 3, aes(label = ..rr.label..)) +
  theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));standard.plot
```

# Calculate surface area using the standard curve  
```{r}
#Calculate surface area using the standard curve
smpls <- subset(wax.data, sample=="sample")
smpls$surface.area.cm2 <- stnd.curve$coefficients[2] * smpls$delta.mass.g + stnd.curve$coefficients[1]

#select the samples only
smpls <- smpls %>%
  select(-sample, -diameter)

smpls %>%
  count(fragment) %>% arrange(n)

#check the range to make sure your samples fall within the range of the standards
range(smpls$surface.area.cm2)
range(stnds$surface.area.cm2)
#yes these fall within the range 
```

# Plot samples on the standard curve.  
```{r}
# Plot standard curve
std_curve_plot <- stnds %>%
  ggplot(aes(x = delta.mass.g, y = surface.area.cm2)) +
  geom_point(color = "red", size = 3) 

std_curve_plot + 
  geom_point(data = smpls, aes(x = delta.mass.g, y = surface.area.cm2), pch = "X", cex = 5, alpha = 0.3) +
  labs(title = "All samples projected on standard curve")
```

All samples are within the standard curve.  

# Save output for normalization for physiology analysis.  

```{r}
#Save the output for use in normilzation for phys assays
smpls%>%
  mutate(timepoint="baseline")%>%
  select(fragment, surface.area.cm2, timepoint)%>%
    write_csv("output/baseline/surface_area_calc.csv")
```
