---
title: Protein analysis
author: "AS Huffmyer"
date: '2022'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---
This script analyzes and plots data for symbiont cell density for the E5 *A. pulchra* metabolism 2022 project. Scripts based on E5 timeseries analysis scripts. 

```{r setup, include = FALSE}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
# load packages
library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
```

# Import data
```{r}
# Cell count data
sym_counts <- read_excel("data/baseline_sampling/sym_counts.xlsx")

# Surface area data
sa <- read.csv("output/baseline/surface_area_calc.csv")

# Tissue homogenate volume data
homog_vols <- read_csv("data/baseline_sampling/homog_vols.csv") %>% select(1:2)

# Coral sample metadata
metadata <- read_csv("data/fragments.csv") %>% select(1:4) %>% filter(treatment=="Baseline")

# Join homogenate volumes and surface area with sample metadata
metadata <- full_join(metadata, homog_vols) %>%
  full_join(sa)
```

# Calculate cells per square centimeter
```{r}
# Calculate mean counts for each sample
sym_counts <- sym_counts %>%
  select(fragment, squares.counted, matches("Count[0-6]")) %>%
  gather("rep", "count", -fragment, -squares.counted) %>%
  group_by(fragment, squares.counted) %>%
  summarise(mean_count = mean(count, na.rm = TRUE))

# Join mean counts with sample metadata
sym_counts <- full_join(sym_counts, metadata)

# Normalize counts by homogenat volume and surface area
sym_counts <- sym_counts %>%
  mutate(cells.mL = mean_count * 10000 / squares.counted,
         cells = cells.mL * homog.vol.mL,
         cells.cm2 = cells / surface.area.cm2)
```

Output file.  

```{r}
sym_counts %>%
  select(fragment, colony, tank, treatment, cells.cm2) %>%
  mutate(timepoint="baseline")%>% 
  write_csv("output/baseline/sym_densities_calc.csv")
```

# Plot data by colony 

Plot by colony  
```{r}
# Plot all data points with mean ± se
sym_plot<-sym_counts %>%
  
  ggplot(aes(x = colony, y = cells.cm2, color = colony)) +
  #coord_cartesian(ylim = c(0, 1.0))+
  labs(x = "", y = "Symbiont Cells per cm2") +
  geom_boxplot()+
  #ylim(200,1000)+
  geom_point(pch = 21, position = position_jitterdodge(0.1))+
  theme_classic(); sym_plot # Plot all points

ggsave(sym_plot, file="figures/baseline/symbiont_density.png", width=6, height=6)
```

These values match the expected range for Acropora from the E5 Timeseries project (~1e+06).  

# Statistical tests  

Analyze by differences in colony using a mixed effect model with tank as a random effect.  

```{r}
model_sym<-sym_counts%>%
  
  lmer(cells.cm2~colony+(1|tank), data=.)

anova(model_sym)
```



