---
title: "Analyzing logger data"
author: "Ariana S Huffmyer"
date: "10/10/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots logger environmental data from Hobo pendant loggers in the E5 *A. pulchra* metabolism experiment.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Loggers: Pendants in E5 tanks  

## Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/E5_tanks", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, temp.C=`Ch: 1 - Temperature   (°C )`, light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 

#adjust time to correct values and force time zone
pendants$DateTime <- force_tz(pendants$DateTime, tzone = "Pacific/Tahiti")
```

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/environmental/loggers/E5_tanks/logger_metadata.csv")
metadata$logger<-as.character(metadata$logger)
```

Assign treatment information to logger data.  
```{r}
pendants<-left_join(pendants, metadata, by="logger")
pendants$tank<-as.factor(pendants$tank)
```

## Apply tempreature calibrations  

Read in calibration output file with the intercept and coefficients for calibration.  
```{r}
temp_cal<-read_csv("output/environmental/temperature_logger_calibrations.csv")
temp_cal$logger<-as.character(temp_cal$logger)
```

Generate a calibrated temperature column in our logger dataframe.  
```{r}
pendants<-left_join(pendants, temp_cal, by="logger")

pendants<-pendants%>%
  mutate(temp.calibrated=temp.C*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Apply light calibrations  

Read in calibration output file with the intercept and coefficients for calibration.  
```{r}
light_cal<-read_csv("output/environmental/light_logger_calibrations.csv")
light_cal$logger<-as.character(light_cal$logger)
```

Generate a calibrated light column in our logger dataframe.  
```{r}
pendants<-left_join(pendants, light_cal, by="logger")

pendants<-pendants%>%
  mutate(PAR.calibrated=light.lux*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Plot temperature    

Plot over time by tank  
```{r}
temp_plot<-pendants%>%
  filter(DateTime > '2022-09-26 11:20:00')%>% 
  
  ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+
  geom_line()+
  theme_classic(); temp_plot
```

Oct 10, 2022: Tanks 1, 2, and 3 have been experiencing a spike in temperature in the morning hours. This is due to the way the sun hits the tanks in the morning from the side. We can move the experiment to a different water table section. This was completed on 10/11/2022.  

## Plot light     

Plot over time by tank  
```{r}
light_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=PAR.calibrated, colour=tank))+
  geom_line()+
  theme_classic(); light_plot
```

Plot by lux (non calibrated)  
```{r}
light_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=light.lux, colour=tank))+
  geom_line()+
  theme_classic(); light_plot
```

Oct 10, 2022: Tanks 1, 2, and 3 have been experiencing a spike in light in the morning hours. This is due to the way the sun hits the tanks in the morning from the side. We can move the experiment to a different water table section. This was completed on 10/11/2022.    
Oct 10, 2022: We need to revisit the light logger calibration. PAR values are higher than expected from the daily measurements.  

# Loggers: Hobo black loggers in parent tanks  

## Read in files  

Read in Hobo black logger files  
```{r}
hobo.files <- list.files(path="data/environmental/loggers/parents", pattern="*.csv", full.names=TRUE)

hobo <- setNames(hobo.files, hobo.files) %>%
   map_dfr(read_csv, .id = "logger", skip=1)%>%
   mutate(logger=sub(".*/", "", logger))%>% #remove the path name
   mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number
   select(!'#')%>%
   select(logger, `Date Time, GMT-10:00`, starts_with("Temp"))%>%
   rename(DateTime=`Date Time, GMT-10:00`, Temp1=`Temp, °C (LGR S/N: 20444033, SEN S/N: 20444033, LBL: C)`, Temp2=`Temp, °C (LGR S/N: 20444040, SEN S/N: 20444040)`, Temp3=`Temp, °C (LGR S/N: 20719656, SEN S/N: 20719656, LBL: C)`, Temp4=`Temp, °C (LGR S/N: 20946644, SEN S/N: 20946644)`)%>%
   mutate(temp.C = coalesce(Temp1, Temp2, Temp3, Temp4))%>%
  select(logger, DateTime, temp.C)

#adjust time to correct values and force time zone
hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")
hobo$DateTime<-as.POSIXct(hobo$DateTime,format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")
hobo$DateTime <- force_tz(hobo$DateTime, tzone = "Pacific/Tahiti")
```

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata_parent<-read_csv("data/environmental/loggers/parents/logger_metadata_parents.csv")
metadata_parent$logger<-as.character(metadata_parent$logger)
```

Assign treatment information to logger data.  
```{r}
hobo<-left_join(hobo, metadata_parent, by="logger")
```

## Apply tempreature calibrations  

Read in calibration output file with the intercept and coefficients for calibration.  
```{r}
temp_cal<-read_csv("output/environmental/temperature_logger_calibrations.csv")
temp_cal$logger<-as.character(temp_cal$logger)
```

Generate a calibrated temperature column in our logger dataframe.  
```{r}
hobo<-left_join(hobo, temp_cal, by="logger")

hobo<-hobo%>%
  mutate(temp.calibrated=temp.C*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Plot temperature    

Plot over time by tank  
```{r}
temp_plot_parents<-hobo%>%
  filter(DateTime < '2022-10-02 11:20:00')%>%
  mutate(temp.calibrated = replace(temp.calibrated, temp.calibrated<25.5, NA))%>%
  
  ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+
  geom_line()+
  theme_classic(); temp_plot_parents
```

# Daily measurements  

## Read in file  

Read in daily measurements file.  

```{r}
daily<-read_csv("data/environmental/daily_measurements/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tank<-as.factor(daily$tank)
daily$tris.date<-as.character(daily$tris.date)
```

## Calculate pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/environmental/20220924_Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~TTris, data = data))) %>% # linear regression of mV and temp of the tris
  summarise(broom::tidy(fitpH)) %>% # make the output tidy
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate)%>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*TTris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  drop_na() %>%
  mutate(pH = pH(Ex=pH.mV,Etris=mVTris,S=sal.psu,T=temp.C)) # calculate pH of the tanks using the pH seacarb function

```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  

```{r}
daily_calc<-pHSlope%>%
  select(date, time, tank, temp.C, pH, sal.psu, flow.mL.5s.final, par.center, par.q1, par.q2, par.q3, par.q4)
```

## Calculate flow  

Calculate flow to total mL per minute rather than 5 sec.  
```{r}
daily_calc<-daily_calc%>%
  mutate(flow.L.min=(flow.mL.5s.final*12)/1000)%>%
  select(!flow.mL.5s.final)
```

## Calculate mean light value 

Calculate the mean light value (PAR) as an average of all light measurements in the 5 positions in each tank at each measurement.  
```{r}
daily_calc<-daily_calc%>%
  mutate(mean.par=rowMeans(across(starts_with("par"))))%>%
  select(date, time, tank, temp.C, pH, sal.psu, flow.L.min, mean.par)
```

## Plot metrics of interest  

Plot temperature values.  
```{r}
daily_temp<-daily_calc%>%
  
  ggplot(aes(x=date, y=temp.C, colour=tank))+
  geom_point()+
  ylim(26.75,28)+
  ylab("Temperature °C")+
  xlab("Date")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic(); daily_temp
  
```

Plot salinity values.  
```{r}
daily_sal<-daily_calc%>%
  
  ggplot(aes(x=date, y=sal.psu, colour=tank))+
  geom_point()+
  ylim(38,38.4)+
  ylab("Salinity (psu)")+
  xlab("Date")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic(); daily_sal
  
```

Plot pH values.  
```{r}
daily_ph<-daily_calc%>%
  
  ggplot(aes(x=date, y=pH, colour=tank))+
  geom_point()+
  ylim(8.03,8.12)+
  ylab("Total pH")+
  xlab("Date")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic(); daily_ph
  
```

Plot flow values.  
```{r}
daily_flow<-daily_calc%>%
  
  ggplot(aes(x=date, y=flow.L.min, colour=tank))+
  geom_point()+
  ylim(3,5.5)+
  ylab("Flow (L per min)")+
  xlab("Date")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic(); daily_flow
```

Plot light PAR values.  
```{r}
daily_par<-daily_calc%>%
  
  ggplot(aes(x=date, y=mean.par, colour=tank))+
  geom_point()+
  ylim(0,600)+
  ylab("Light (PAR)")+
  xlab("Date")+
  scale_fill_brewer(palette = "Dark2")+
  theme_classic(); daily_par
```

View a panel of all plots together.  
```{r}
daily_temp_l<-daily_temp+theme(legend.position="none")
daily_ph_l<-daily_ph+theme(legend.position="none")
daily_sal_l<-daily_sal+theme(legend.position="none")
daily_par_l<-daily_par+theme(legend.position="none")
daily_flow_l<-daily_flow+theme(legend.position="none")

daily_plots<-plot_grid(daily_temp_l, daily_ph_l, daily_sal_l, daily_par_l, daily_flow, nrow=1, ncol=5)  

ggsave(filename="figures/environmental/daily_measurements.png", plot=daily_plots, dpi=300, width=24, height=5, units="in")
```

## Test for differences between tanks  

Test for differences in temperature between tanks.  
```{r}
summary(aov(temp.C~tank, data=daily_calc))
```

Test for differences in pH between tanks.  
```{r}
summary(aov(pH~tank, data=daily_calc))
```

Test for differences in salinity between tanks.  
```{r}
summary(aov(sal.psu~tank, data=daily_calc))
```

Test for differences in flow between tanks.  
```{r}
summary(aov(flow.L.min~tank, data=daily_calc))
```

Test for differences in light between tanks.  
```{r}
summary(aov(mean.par~tank, data=daily_calc))
```