---
title: "Analyzing logger data"
author: "Ariana S Huffmyer"
date: "10/10/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots logger environmental data from Hobo pendant loggers in the E5 *A. pulchra* metabolism experiment.  

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Loggers: Pendants in E5 tanks  

## Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/E5_tanks", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, temp.C=`Ch: 1 - Temperature   (°C )`, light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 

#adjust time to correct values and force time zone
pendants$DateTime <- force_tz(pendants$DateTime, tzone = "Pacific/Tahiti")
```

## Read in treatment information  

Read in metadata that assigns a pendant serial number to a tank number and/or treatment.  
```{r}
metadata<-read_csv("data/environmental/loggers/E5_tanks/logger_metadata.csv")
metadata$logger<-as.character(metadata$logger)
```

Assign treatment information to logger data.  
```{r}
pendants<-left_join(pendants, metadata, by="logger")
pendants$tank<-as.factor(pendants$tank)
```

## Apply tempreature calibrations  

Read in calibration output file with the intercept and coefficients for calibration.  
```{r}
temp_cal<-read_csv("output/environmental/temperature_logger_calibrations.csv")
temp_cal$logger<-as.character(temp_cal$logger)
```

Generate a calibrated temperature column in our logger dataframe.  
```{r}
pendants<-left_join(pendants, temp_cal, by="logger")

pendants<-pendants%>%
  mutate(temp.calibrated=temp.C*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Apply light calibrations  

Read in calibration output file with the intercept and coefficients for calibration.  
```{r}
light_cal<-read_csv("output/environmental/light_logger_calibrations.csv")
light_cal$logger<-as.character(light_cal$logger)
```

Generate a calibrated light column in our logger dataframe.  
```{r}
pendants<-left_join(pendants, light_cal, by="logger")

pendants<-pendants%>%
  mutate(PAR.calibrated=light.lux*coef+intercept)%>%
  select(!coef)%>%
  select(!intercept)
```

## Plot temperature    

Plot over time by tank  
```{r}
temp_plot<-pendants%>%
  filter(DateTime > '2022-09-26 11:20:00')%>% 
  
  ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+
  geom_line()+
  theme_classic(); temp_plot
```

Oct 10, 2022: Tanks 1, 2, and 3 have been experiencing a spike in temperature in the morning hours. This is due to the way the sun hits the tanks in the morning from the side. We can move the experiment to a different water table section. This was completed on 10/11/2022.  

## Plot light     

Plot over time by tank  
```{r}
light_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=PAR.calibrated, colour=tank))+
  ylab("Integrated PAR")+
  geom_line()+
  theme_classic(); light_plot
```

Plot by lux (non calibrated)  
```{r}
lux_plot<-pendants%>%
  ggplot(aes(x=DateTime, y=light.lux, colour=tank))+
  ylab("Light Lux")+
  geom_line()+
  theme_classic(); lux_plot
```

Oct 10, 2022: Tanks 1, 2, and 3 have been experiencing a spike in light in the morning hours. This is due to the way the sun hits the tanks in the morning from the side. We can move the experiment to a different water table section. This was completed on 10/11/2022.    
Oct 10, 2022: We need to revisit the light logger calibration. PAR values are higher than expected from the daily measurements.  
<!-- # Loggers: Pendants in larval tanks     -->

<!-- ## Read in files   -->

<!-- Read in Hobo Pendant files   -->
<!-- ```{r} -->
<!-- pendant.files.larvae <- list.files(path="data/environmental/loggers/larvae", pattern="*.xlsx", full.names=TRUE) -->

<!-- pendants.larvae <- setNames(pendant.files.larvae, pendant.files.larvae) %>% -->
<!--    map_dfr(read_excel, .id = "logger") %>% -->
<!--    rename(DateTime=`Date-Time (French Polynesia Standard Time)`, temp.C=`Ch: 1 - Temperature   (°C )`, light.lux=`Ch: 2 - Light   (lux)`)%>% -->
<!--   select(!`#`)%>% -->
<!--   mutate(logger=sub(".*/", "", logger))%>% #remove the path name -->
<!--   mutate(logger=sub("_.*", "", logger)) #keep only the serial number  -->

<!-- #adjust time to correct values and force time zone -->
<!-- pendants.larvae$DateTime <- force_tz(pendants.larvae$DateTime, tzone = "Pacific/Tahiti") -->
<!-- ``` -->

<!-- ## Read in treatment information   -->

<!-- Read in metadata that assigns a pendant serial number to a tank number and/or treatment.   -->
<!-- ```{r} -->
<!-- metadata_larvae<-read_csv("data/environmental/loggers/larvae/logger_metadata_larvae.csv") -->
<!-- metadata_larvae$logger<-as.character(metadata_larvae$logger) -->
<!-- ``` -->

<!-- Assign treatment information to logger data.   -->
<!-- ```{r} -->
<!-- pendants.larvae<-left_join(pendants.larvae, metadata_larvae, by="logger") -->
<!-- pendants.larvae$tank<-as.factor(pendants.larvae$tank) -->
<!-- ``` -->

<!-- ## Apply tempreature calibrations   -->

<!-- Read in calibration output file with the intercept and coefficients for calibration.   -->
<!-- ```{r} -->
<!-- temp_cal<-read_csv("output/environmental/temperature_logger_calibrations.csv") -->
<!-- temp_cal$logger<-as.character(temp_cal$logger) -->
<!-- ``` -->

<!-- Generate a calibrated temperature column in our logger dataframe.   -->
<!-- ```{r} -->
<!-- pendants.larvae<-left_join(pendants.larvae, temp_cal, by="logger") -->

<!-- pendants.larvae<-pendants.larvae%>% -->
<!--   mutate(temp.calibrated=temp.C*coef+intercept)%>% -->
<!--   select(!coef)%>% -->
<!--   select(!intercept) -->
<!-- ``` -->

<!-- ## Apply light calibrations   -->

<!-- Read in calibration output file with the intercept and coefficients for calibration.   -->
<!-- ```{r} -->
<!-- light_cal<-read_csv("output/environmental/light_logger_calibrations.csv") -->
<!-- light_cal$logger<-as.character(light_cal$logger) -->
<!-- ``` -->

<!-- Generate a calibrated light column in our logger dataframe.   -->
<!-- ```{r} -->
<!-- pendants.larvae<-left_join(pendants.larvae, light_cal, by="logger") -->

<!-- pendants.larvae<-pendants.larvae%>% -->
<!--   mutate(PAR.calibrated=light.lux*coef+intercept)%>% -->
<!--   select(!coef)%>% -->
<!--   select(!intercept) -->
<!-- ``` -->

<!-- ## Plot temperature     -->

<!-- Plot over time by tank   -->
<!-- ```{r} -->
<!-- temp_plot_larvae<-pendants.larvae%>% -->

<!--   ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+ -->
<!--   geom_line()+ -->
<!--   theme_classic(); temp_plot_larvae -->
<!-- ``` -->

<!-- Oct 10, 2022: Tanks 1, 2, and 3 have been experiencing a spike in temperature in the morning hours. This is due to the way the sun hits the tanks in the morning from the side. We can move the experiment to a different water table section. This was completed on 10/11/2022.   -->

<!-- ## Plot light      -->

<!-- Plot over time by tank   -->
<!-- ```{r} -->
<!-- light_plot_larvae<-pendants.larvae%>% -->
<!--   ggplot(aes(x=DateTime, y=PAR.calibrated, colour=tank))+ -->
<!--   ylab("Integrated PAR")+ -->
<!--   geom_line()+ -->
<!--   theme_classic(); light_plot_larvae -->
<!-- ``` -->

<!-- Plot by lux (non calibrated)   -->
<!-- ```{r} -->
<!-- lux_plot_larvae<-pendants.larvae%>% -->
<!--   ggplot(aes(x=DateTime, y=light.lux, colour=tank))+ -->
<!--   ylab("Light Lux")+ -->
<!--   geom_line()+ -->
<!--   theme_classic(); lux_plot_larvae -->
<!-- ``` -->



<!-- # Loggers: Hobo black loggers in parent tanks   -->

<!-- ## Read in files   -->

<!-- Read in Hobo black logger files   -->
<!-- ```{r} -->
<!-- hobo.files <- list.files(path="data/environmental/loggers/parents", pattern="*.csv", full.names=TRUE) -->

<!-- hobo <- setNames(hobo.files, hobo.files) %>% -->
<!--    map_dfr(read_csv, .id = "logger", skip=1)%>% -->
<!--    mutate(logger=sub(".*/", "", logger))%>% #remove the path name -->
<!--    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number -->
<!--    select(!'#')%>% -->
<!--    select(logger, `Date Time, GMT-10:00`, starts_with("Temp"))%>% -->
<!--    rename(DateTime=`Date Time, GMT-10:00`, Temp1=`Temp, °C (LGR S/N: 20444033, SEN S/N: 20444033, LBL: C)`, Temp2=`Temp, °C (LGR S/N: 20444040, SEN S/N: 20444040)`, Temp3=`Temp, °C (LGR S/N: 20719656, SEN S/N: 20719656, LBL: C)`, Temp4=`Temp, °C (LGR S/N: 20946644, SEN S/N: 20946644)`)%>% -->
<!--    mutate(temp.C = coalesce(Temp1, Temp2, Temp3, Temp4))%>% -->
<!--   select(logger, DateTime, temp.C) -->

<!-- #adjust time to correct values and force time zone -->
<!-- hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti") -->
<!-- hobo$DateTime<-as.POSIXct(hobo$DateTime,format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti") -->
<!-- hobo$DateTime <- force_tz(hobo$DateTime, tzone = "Pacific/Tahiti") -->
<!-- ``` -->

<!-- ## Read in treatment information   -->

<!-- Read in metadata that assigns a pendant serial number to a tank number and/or treatment.   -->
<!-- ```{r} -->
<!-- metadata_parent<-read_csv("data/environmental/loggers/parents/logger_metadata_parents.csv") -->
<!-- metadata_parent$logger<-as.character(metadata_parent$logger) -->
<!-- ``` -->

<!-- Assign treatment information to logger data.   -->
<!-- ```{r} -->
<!-- hobo<-left_join(hobo, metadata_parent, by="logger") -->
<!-- ``` -->

<!-- ## Apply tempreature calibrations   -->

<!-- Read in calibration output file with the intercept and coefficients for calibration.   -->
<!-- ```{r} -->
<!-- temp_cal<-read_csv("output/environmental/temperature_logger_calibrations.csv") -->
<!-- temp_cal$logger<-as.character(temp_cal$logger) -->
<!-- ``` -->

<!-- Generate a calibrated temperature column in our logger dataframe.   -->
<!-- ```{r} -->
<!-- hobo<-left_join(hobo, temp_cal, by="logger") -->

<!-- hobo<-hobo%>% -->
<!--   mutate(temp.calibrated=temp.C*coef+intercept)%>% -->
<!--   select(!coef)%>% -->
<!--   select(!intercept) -->
<!-- ``` -->

<!-- ## Plot temperature     -->

<!-- Plot over time by tank   -->
<!-- ```{r} -->
<!-- temp_plot_parents<-hobo%>% -->
<!--   #filter(DateTime < '2022-10-0 11:20:00')%>% -->
<!--   mutate(temp.calibrated = replace(temp.calibrated, temp.calibrated<25.5, NA))%>% -->

<!--   ggplot(aes(x=DateTime, y=temp.calibrated, colour=tank))+ -->
<!--   geom_line()+ -->
<!--   theme_classic(); temp_plot_parents -->
<!-- ``` -->

<!-- # Generate figure of all logger data   -->

<!-- temp_plot -->
<!-- light_plot -->
<!-- lux_plot -->
<!-- temp_plot_parents -->

<!-- temp_plot_larvae -->
<!-- light_plot_larvae -->
<!-- lux_plot_larvae -->

<!-- ```{r} -->
<!-- temp_plot_l<-temp_plot+theme(legend.position="none") -->
<!-- light_plot_l<-light_plot+theme(legend.position="none") -->

<!-- logger_plots<-plot_grid(temp_plot, light_plot, lux_plot, temp_plot_parents, temp_plot_larvae, light_plot_larvae, lux_plot_larvae, nrow=7, ncol=1)   -->

<!-- ggsave(filename="figures/environmental/loggers.png", plot=logger_plots, dpi=300, width=9, height=30, units="in") -->

<!-- ``` -->

# Daily measurements  

## Read in file  

Read in daily measurements file.  

```{r}
daily<-read_csv("data/environmental/daily_measurements/daily_measurements.csv")
daily$date<-as.Date(daily$date, format="%m/%d/%y")
daily$tank<-as.factor(daily$tank)
daily$tris.date<-as.character(daily$tris.date)
```

## Calculate pH  

Calculate the calibration curve from the Tris calibration and calculate pH on the total scale from pH.mV.   
```{r}
pHcalib<-read_csv("data/environmental/20220924_Tris_Calibration.csv")
pHcalib$tris.date<-as.character(pHcalib$tris.date)

pHSlope<-pHcalib %>%
  nest_by(tris.date)%>%
  mutate(fitpH = list(lm(mVTris~TTris, data = pHcalib))) %>% # linear regression of mV and temp of the tris
  summarise(broom::tidy(fitpH)) %>% # make the output tidy
  select(tris.date, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%# put slope and intercept in their own column
  left_join(daily, ., by="tris.date") %>% # join with the pH sample data
  mutate(mVTris = temp.C*TTris + `(Intercept)`) %>%# calculate the mV of the tris at temperature in which the pH of the tanks were measured
  mutate(pH = pH(Ex=pH.mV,Etris=mVTris,S=sal.psu,T=temp.C)) # calculate pH of the tanks using the pH seacarb function

```

pH is now calculated as Total pH in the "pH" column. Now select the desired columns to analyze further.  

```{r}
daily_calc<-pHSlope%>%
  select(date, time, tank, treatment, Period, temp.C, pH, sal.psu, flow.mL.5s.initial, par.center)
```

## Calculate flow  

Calculate flow to total mL per minute rather than 5 sec.  
```{r}
daily_calc<-daily_calc%>%
  mutate(flow.L.min=(flow.mL.5s.initial*12)/1000)%>%
  select(!flow.mL.5s.initial)
```

## Calculate mean light value 

Calculate the mean light value (PAR) as an average of all light measurements in the 5 positions in each tank at each measurement.  
```{r}
# daily_calc<-daily_calc%>%
#   mutate(mean.par=rowMeans(across(starts_with("par")), na.rm=TRUE)) %>%
#   select(date, time, tank, temp.C, pH, sal.psu, flow.L.min, par.center, mean.par)

daily_calc$par.center <- as.numeric(as.character(daily_calc$par.center))
```

change data format to long format 
```{r}
daily_calc.long <-daily_calc %>% pivot_longer(cols=temp.C:flow.L.min,
  names_to = "metric",
  values_to = "value")
```

## Plot metrics of interest  
Plot by tank.  
```{r}
daily_tanks<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=tank))+
  geom_point()+
  geom_vline(xintercept = as.numeric(ymd("2022-10-18")), linetype="dashed", 
                color = "black", size=0.5)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  scale_fill_brewer(palette = "Dark2")+
  theme_bw(); daily_tanks
  
ggsave(filename="figures/environmental/tank_daily_measurements.png", plot=daily_tanks, dpi=300, width=8, height=10, units="in")

```

Plot by treatment  
```{r}
daily_plots<-daily_calc.long %>%
  ggplot(aes(x=date, y=value, colour=treatment))+
  geom_point()+
  geom_vline(xintercept = as.numeric(ymd("2022-10-18")), linetype="dashed", 
                color = "black", size=0.5)+
  xlab("Date")+
  facet_grid(metric ~ ., scales = "free")+
  scale_fill_brewer(palette = "Dark2")+
  theme_bw(); daily_plots

ggsave(filename="figures/environmental/daily_measurements.png", plot=daily_plots, dpi=300, width=8, height=10, units="in")
```

Separate into Acclimation period and Exposure Period
```{r}
exposure.data <- daily_calc.long %>%
  filter(Period=="Exposure")

```

Plot by treatment  
```{r}
daily_plots<-exposure.data %>%
  ggplot(aes(x=treatment, y=value, fill=treatment))+
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  geom_vline(xintercept = as.numeric(ymd("2022-10-18")), linetype="dashed", 
                color = "black", size=0.5)+
  xlab("Treatment")+
  facet_grid(metric ~ ., scales = "free")+
  scale_fill_brewer(palette = "Dark2")+
  theme_bw(); daily_plots

ggsave(filename="figures/environmental/treatment_daily_measurements.png", plot=daily_plots, dpi=300, width=8, height=10, units="in")
```


## Test for differences between tanks  
```{r}

daily_calc_exposure <- daily_calc%>%
  filter(Period=="Exposure")

#Temp
summary(aov(temp.C~tank, data=daily_calc_exposure))

#pH
summary(aov(pH~tank, data=daily_calc_exposure))

#Salinity
summary(aov(sal.psu~tank, data=daily_calc_exposure))

#Flow
summary(aov(flow.L.min~tank, data=daily_calc_exposure))

#Light
summary(aov(par.center~tank, data=daily_calc_exposure))
```

## Test for differences between Treatments  
```{r}

daily_calc_exposure <- daily_calc%>%
  filter(Period=="Exposure")

#Temp
summary(aov(temp.C~treatment, data=daily_calc_exposure))

#pH
summary(aov(pH~treatment, data=daily_calc_exposure))

#Salinity
summary(aov(sal.psu~treatment, data=daily_calc_exposure))

#Flow
summary(aov(flow.L.min~treatment, data=daily_calc_exposure))

#Light
summary(aov(par.center~treatment, data=daily_calc_exposure))
```
