---
title: "Calibrating light and temperature loggers"
author: "Ariana S Huffmyer"
date: "9/30/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
```

# Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/calibration/pendants", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, Temp.C.pendant=`Ch: 1 - Temperature   (°C )`, Light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 

#adjust time to correct values and force time zone
pendants$DateTime <- force_tz(pendants$DateTime, tzone = "Pacific/Tahiti")
```

Read in Hobo temperature files.     
```{r}
hobo.files <- list.files(path="data/environmental/loggers/calibration/hobo", pattern="*.csv", full.names=TRUE)

hobo <- setNames(hobo.files, hobo.files) %>%
   map_dfr(read_csv, .id = "logger", skip=1)%>%
   mutate(logger=sub(".*/", "", logger))%>% #remove the path name
   mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number
   select(!'#')%>%
   select(logger, `Date Time, GMT-10:00`, starts_with("Temp"))%>%
   rename(DateTime=`Date Time, GMT-10:00`, Temp1=`Temp, °C (LGR S/N: 20444033, SEN S/N: 20444033, LBL: C)`, Temp2=`Temp, °C (LGR S/N: 20719656, SEN S/N: 20719656, LBL: C)`, Temp3=`Temp, °C (LGR S/N: 20444040, SEN S/N: 20444040)`, Temp4=`Temp, °C (LGR S/N: 20946644, SEN S/N: 20946644)`)%>%
   mutate(Temp.C.hobo = coalesce(Temp1, Temp2, Temp3, Temp4))%>%
  select(logger, DateTime, Temp.C.hobo)

#convert Date Time to posix 
hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")
hobo$DateTime<-as.POSIXct(hobo$DateTime,format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")
```

Read in Odyssey light logger.  
```{r}
light<-read_csv("data/environmental/loggers/calibration/odyssey/15643_20220925_light_calib.csv", skip=7)

colnames(light)<-c("number", "Date", "Time", "Raw.PAR", "Calibrated.PAR")

light <- light %>%
  mutate(DateTime = paste(Date, Time))%>%
  mutate(DateTime = as.POSIXct(DateTime, format="%d/%m/%Y %H:%M:%S", tz="Pacific/Tahiti"))%>%
  select(DateTime, Calibrated.PAR)%>%
  rename(Light.PAR=Calibrated.PAR)%>%
  mutate(logger="15643")
```

# Merge files together  

First make a list of the SNs for each type of logger before spreading to wide format for each.  
```{r}
hobo_list<-levels(as.factor(hobo$logger))
pendant_list<-levels(as.factor(pendants$logger))
light_list<-levels(as.factor(light$logger))
```

Convert each data frame to wide format.  
```{r}
hobo<-hobo%>%
  spread(value=Temp.C.hobo, key=logger)

#make two for pendants - one for light one for temp
pendants_light<-pendants%>%
  select(!Temp.C.pendant)%>%
  spread(value=Light.lux, key=logger)

pendants_temp<-pendants%>%
  select(!Light.lux)%>%
  spread(value=Temp.C.pendant, key=logger)

light<-light%>%
  spread(value=Light.PAR, key=logger)
```

Confirm time zones are the same.  

```{r}
attr(light$DateTime,"tzone")
attr(hobo$DateTime,"tzone")
attr(pendants$DateTime,"tzone")
```

Merge data frames.  

```{r}
cal_light<-inner_join(pendants_light, light)
cal_temp<-inner_join(hobo, pendants_temp)
```

Remove time periods when the loggers were being read out or launched.  
```{r}
cal_light <- cal_light %>% 
  filter(DateTime > '2022-09-23 17:20:00')%>%
  filter(DateTime < '2022-09-25 11:00:00')

cal_temp <- cal_temp %>% 
  filter(DateTime > '2022-09-23 17:20:00')%>%
  filter(DateTime < '2022-09-25 11:00:00')
```


# Temperature calibrations 

## 1. Plot temperature logging over time for each logger and color by black loggers and pendants.  

```{r}
temp_plot<-cal_temp%>%
  gather(value=temp, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=temp, colour=logger))+
  geom_line()+
  theme_classic(); temp_plot

```

## 2. Calculate mean hobo loggers, this will be our standard temperature that we will calibrate all else to.    

Calculate mean hobo temperature values.  
```{r}
cal_temp<-cal_temp%>%
  mutate(mean_hobo=rowMeans(across(hobo_list)))
```

## 3. Calculate correlation curves for all loggers  

Run models for each logger against the standard.  
```{r}
df_temp_cor<-cal_temp%>%
  select(!DateTime)

temp_output<-purrr::map(

  # set the elements to iterate over.
  # in this case all variables except 
  # the first (mpg)
  df_temp_cor[,-23],

  # The second argument is the body of the loop. 
  # An lm call with the formula as follow.
  # here the .x is replace by a new variable 
  # in each iteration
  ~lm(mean_hobo ~ .x, data = df_temp_cor)
  ) %>%


  # Then summarise each output with broom::glance
  purrr::map(broom::glance) %>%

  # bind all summary
  dplyr::bind_rows(.id = "variable") %>%

  # selecting the variables of interest
  dplyr::select(variable, r.squared)

temp_output
```

This table shows that r squared fit of each logger to the mean Hobo logger temperatures are 0.89-0.99. 









## 4. Output to a data frame to use in future analysis with these loggers.  







# Light calibrations  

## 1. Plot temperature logging over time for each logger and color by odyssey and pendants.  
```{r}
light_plot<-cal_light%>%
  #select(DateTime, `15643`)%>%
  gather(value=light, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=light, colour=logger))+
  geom_line()+
  theme_classic(); light_plot
```

## 2. Calculate the correlation of each logger to the mean hobo loggers for each individual logger. 

```{r}
df_light_cor<-cal_light%>%
  select(!DateTime)

light_cor <- cor(df_light_cor[ , colnames(df_light_cor) != "15643"], df_light_cor$`15643`, use = "na.or.complete")
light_cor  
```

Correlation values to the Odyssey logger range from 0.88 to 0.95. Due to small scale variation in light values in the tank, these values are appropriate.  








Calculate the correlation curve at each observation from the odyssey loggers for each individual logger.  

Output to a data frame to use in future analysis with these loggers.  