---
title: "Calibrating light and temperature loggers"
author: "Ariana S Huffmyer"
date: "9/30/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(PerformanceAnalytics)
```


# Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/calibration/pendants", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, Temp.C.pendant=`Ch: 1 - Temperature   (°C )`, Light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 

pendants$DateTime<-as.POSIXct(pendants$DateTime, format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")

#adjust time to correct values and force time zone
pendants$DateTime <- force_tz(pendants$DateTime, tzone = "Pacific/Tahiti")
```

Read in Hobo temperature files.     
```{r}
hobo.files <- list.files(path="data/environmental/loggers/calibration/hobo", pattern="*.csv", full.names=TRUE)

hobo <- setNames(hobo.files, hobo.files) %>%
   map_dfr(read_csv, .id = "logger", skip=1)%>%
   mutate(logger=sub(".*/", "", logger))%>% #remove the path name
   mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number
   select(!'#')%>%
   select(logger, `Date Time, GMT-10:00`, starts_with("Temp"))%>%
   rename(DateTime=`Date Time, GMT-10:00`, Temp1=`Temp, °C (LGR S/N: 20444033, SEN S/N: 20444033, LBL: C)`, Temp2=`Temp, °C (LGR S/N: 20719656, SEN S/N: 20719656, LBL: C)`, Temp3=`Temp, °C (LGR S/N: 20444040, SEN S/N: 20444040)`, Temp4=`Temp, °C (LGR S/N: 20946644, SEN S/N: 20946644)`)%>%
   mutate(Temp.C.hobo = coalesce(Temp1, Temp2, Temp3, Temp4))%>%
  select(logger, DateTime, Temp.C.hobo)

#convert Date Time to posix 
hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S")

#hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")

hobo$DateTime<-as.POSIXct(hobo$DateTime,format="%m/%d/%y %H:%M:%S", tz="Pacific/Tahiti")
```

Read in Odyssey light logger.  
```{r}
light<-read_csv("data/environmental/loggers/calibration/odyssey/15643_20220925_light_calib.csv", skip=7)

colnames(light)<-c("number", "Date", "Time", "Raw.PAR", "Calibrated.PAR")

light <- light %>%
  mutate(DateTime = paste(Date, Time))%>%
  mutate(DateTime = as.POSIXct(DateTime, format="%d/%m/%Y %H:%M:%S", tz="Pacific/Tahiti"))%>%
  select(DateTime, Calibrated.PAR)%>%
  rename(Light.PAR=Calibrated.PAR)%>%
  mutate(logger="15643")
```

# Merge files together  

First make a list of the SNs for each type of logger before spreading to wide format for each.  
```{r}
hobo_list<-levels(as.factor(hobo$logger))
pendant_list<-levels(as.factor(pendants$logger))
light_list<-levels(as.factor(light$logger))
```

Convert each data frame to wide format.  
```{r}
hobo<-hobo%>%
  spread(value=Temp.C.hobo, key=logger)

#make two for pendants - one for light one for temp
pendants_light<-pendants%>%
  filter(Light.lux<40000)%>% #remove outlier points
  select(!Temp.C.pendant)%>%
  spread(value=Light.lux, key=logger)

pendants_temp<-pendants%>%
  select(!Light.lux)%>%
  spread(value=Temp.C.pendant, key=logger)

light<-light%>%
  spread(value=Light.PAR, key=logger)
```

Confirm time zones are the same.  

```{r}
attr(light$DateTime,"tzone")
attr(hobo$DateTime,"tzone")
attr(pendants$DateTime,"tzone")
```

Merge data frames.  

```{r}
cal_light<-inner_join(pendants_light, light)
cal_temp<-inner_join(hobo, pendants_temp)
```

Remove time periods when the loggers were being read out or launched.  

```{r}
cal_light <- cal_light %>% 
  filter(DateTime > as.POSIXct("2022-09-23 17:20:00", tz="Pacific/Tahiti"))%>%
  filter(DateTime < as.POSIXct("2022-09-25 11:00:00", tz="Pacific/Tahiti"))

cal_temp <- cal_temp %>% 
  filter(DateTime > as.POSIXct('2022-09-23 17:20:00', tz="Pacific/Tahiti"))%>%
  filter(DateTime < as.POSIXct('2022-09-25 11:00:00', tz="Pacific/Tahiti"))
```

# Temperature calibrations 

## 1. Plot temperature logging over time for each logger and color by black loggers and pendants.  

```{r}
temp_plot<-cal_temp%>%
  gather(value=temp, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=temp, colour=logger))+
  geom_line()+
  theme_classic(); temp_plot

```

## 2. Calculate mean hobo loggers, this will be our standard temperature that we will calibrate all else to.    

Calculate mean hobo temperature values.  
```{r}
cal_temp<-cal_temp%>%
  mutate(mean_hobo=rowMeans(across(hobo_list)))
```

## 3. Calculate correlation curves for all loggers  

Run models for each logger against the standard.  
```{r}
df_temp_cor<-cal_temp%>%
  select(!DateTime)

temp_output<-purrr::map(

  # set the elements to iterate over.
  # in this case all variables except 
  # the first (mpg)
  df_temp_cor[,-23],

  # The second argument is the body of the loop. 
  # An lm call with the formula as follow.
  # here the .x is replace by a new variable 
  # in each iteration
  ~lm(mean_hobo ~ .x, data = df_temp_cor)
  ) %>%


   # Then summarise each output with broom
  purrr::map(broom::tidy) %>%

  # bind all summary
  dplyr::bind_rows(.id = "variable") %>%

  # selecting the variables of interest
  dplyr::select(variable, term, estimate) %>%

  #Generate columns for intercept and coef for each logger  
  tidyr::spread(value="estimate",key="term")%>%
  rename(coef=".x")%>%
  rename(intercept=`(Intercept)`)%>%
  rename(logger="variable")

temp_output
```

View the results.  
```{r}
temp_output
```

Output to a data frame to use in future analysis with these loggers with the serial number, the coef, and the intercept.    
```{r}
write_csv(temp_output, "output/environmental/temperature_logger_calibrations.csv")
```

View plots of each logger to the mean Hobo logger.    

```{r}
df_temp_cor %>%
  gather(-mean_hobo, key = "var", value = "value") %>% 
  ggplot(aes(x = value, y = mean_hobo)) +
    facet_wrap(~ var, scales = "free") +
    geom_point() +
    ylab("Hobo Mean Logger C")+
    xlab("Pendant Logger C")+
    stat_smooth(method="lm")
```

The formula for calculating calibrations will be:  

(Logger value in C * coef) + intercept = calibrated value in C

# Light calibrations  

## 1. Plot temperature logging over time for each logger and color by odyssey and pendants.  
```{r}
light_plot<-cal_light%>%
  #select(DateTime, `15643`)%>%
  gather(value=light, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=light, colour=logger))+
  geom_line()+
  
  theme_classic(); light_plot
```

Plot only the Odyssey logger.  
```{r}
odyssey_plot<-cal_light%>%
  gather(value=light, key=logger, -DateTime)%>%
  filter(logger=="15643")%>%
  
  ggplot(aes(x=DateTime, y=light, colour=logger))+
  geom_line()+
  theme_classic(); odyssey_plot
```

Plot only the pendant loggers.  
```{r}
pendant_light_plot<-cal_light%>%
  gather(value=light, key=logger, -DateTime)%>%
  filter(!logger=="15643")%>%
  
  ggplot(aes(x=DateTime, y=light, colour=logger))+
  geom_line()+
  theme_classic(); pendant_light_plot
```

## 2. Calculate mean pendant Lux value

This will be our mean Lux value that we will calibrate all else to using an offset calculation using an offset method.    

The Odyssey logger was logging over intervals of 10 minutes. Since the pendants are reading in instantaneous Lux, we are going to first calibrate the pendant loggers Lux values to the mean value of all pendant loggers as we did above for pendant temperature values. We will then apply a conversion factor (0.0185) to convert Lux to PPFD (PAR, umol m2 s) as calculated by Apogee for sunlight. https://www.apogeeinstruments.com/conversion-ppfd-to-lux/   

Calculate mean pendant light values.  
```{r}
cal_light<-cal_light%>%
  select(!'15643')%>%
  mutate(mean_lux=rowMeans(across(pendant_list), na.rm=TRUE))
```

## 3. Calculate correlation curves for all loggers  

Run models for each logger against the group mean.    
```{r}
df_light_cor<-cal_light%>%
  select(!DateTime)

light_output<-purrr::map(

  # set the elements to iterate over.
  # in this case all variables except 
  # the first (mpg)
  df_light_cor[,-19],

  # The second argument is the body of the loop. 
  # An lm call with the formula as follow.
  # here the .x is replace by a new variable 
  # in each iteration
  ~lm(mean_lux ~ .x, data = df_light_cor)
  ) %>%


   # Then summarise each output with broom
  purrr::map(broom::tidy) %>%

  # bind all summary
  dplyr::bind_rows(.id = "variable") %>%

  # selecting the variables of interest
  dplyr::select(variable, term, estimate) %>%

  #Generate columns for intercept and coef for each logger  
  tidyr::spread(value="estimate",key="term")%>%
  rename(coef=".x")%>%
  rename(intercept=`(Intercept)`)%>%
  rename(logger="variable")

```

View the results.  
```{r}
light_output
```

Output to a data frame to use in future analysis with these loggers with the serial number, the coef, and the intercept.    
```{r}
write_csv(light_output, "output/environmental/light_logger_calibrations.csv")
```

View plots of each logger to the Odyssey logger.    

```{r}
df_light_cor %>%
  gather(-mean_lux, key = "var", value = "value") %>% 
  ggplot(aes(x = mean_lux, y = value)) +
    facet_wrap(~ var, scales = "free") +
    geom_point() +
    xlab("Pendant Group Mean")+
    ylab("Pendant Logger Lux")+
    stat_smooth(method="lm")
```

The formula for calculating calibrations will be:  

(Logger value in lux * coef) + intercept = calibrated value  

## 4. Finally, apply the Lux to PAR conversion factor to view light values in PAR 

This will need to be applied to scripts analyzing light data after applying the lux calibration value.  

```{r}
df_light_cor %>%
  gather(-mean_lux, key = "var", value = "value") %>% 
  mutate(par=value*0.0185)%>%
  
  ggplot(aes(x = value, y = par)) +
    facet_wrap(~ var, scales = "free") +
    geom_point() +
    ylab("PAR")+
    xlab("LuxLux")+
    stat_smooth(method="lm")
```

Maximum PAR values were 400-600 which is the range expected from the tank systems.  

View a time series of light data with PAR values.  

Plot only the pendant loggers.  
```{r}
pendant_par_plot<-cal_light%>%
  gather(-DateTime, key = "var", value = "value") %>% 
  filter(!var=="15643")%>%
  mutate(par=value*0.0185)%>%
  #select(!value)%>%
  #spread(key="var", value="par")%>%

  ggplot(aes(x=DateTime, y=par, colour=var))+
  geom_line()+
  theme_classic(); pendant_par_plot
```




