---
title: "Calibrating light and temperature loggers"
author: "Ariana S Huffmyer"
date: "9/30/2022"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
```

# Read in files  

Read in Hobo Pendant files  
```{r}
pendant.files <- list.files(path="data/environmental/loggers/calibration/pendants", pattern="*.xlsx", full.names=TRUE)

pendants <- setNames(pendant.files, pendant.files) %>%
   map_dfr(read_excel, .id = "logger") %>%
   rename(DateTime=`Date-Time (French Polynesia Standard Time)`, Temp.C.pendant=`Ch: 1 - Temperature   (°C )`, Light.lux=`Ch: 2 - Light   (lux)`)%>%
  select(!`#`)%>%
  mutate(logger=sub(".*/", "", logger))%>% #remove the path name
  mutate(logger=sub("_.*", "", logger)) #keep only the serial number 
```

Read in Hobo temperature files.     
```{r}
hobo.files <- list.files(path="data/environmental/loggers/calibration/hobo", pattern="*.csv", full.names=TRUE)

hobo <- setNames(hobo.files, hobo.files) %>%
   map_dfr(read_csv, .id = "logger", skip=1)%>%
   mutate(logger=sub(".*/", "", logger))%>% #remove the path name
   mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number
   select(!'#')%>%
   select(logger, `Date Time, GMT-10:00`, starts_with("Temp"))%>%
   rename(DateTime=`Date Time, GMT-10:00`, Temp1=`Temp, °C (LGR S/N: 20444033, SEN S/N: 20444033, LBL: C)`, Temp2=`Temp, °C (LGR S/N: 20719656, SEN S/N: 20719656, LBL: C)`, Temp3=`Temp, °C (LGR S/N: 20444040, SEN S/N: 20444040)`, Temp4=`Temp, °C (LGR S/N: 20946644, SEN S/N: 20946644)`)%>%
   mutate(Temp.C.hobo = coalesce(Temp1, Temp2, Temp3, Temp4))%>%
  select(logger, DateTime, Temp.C.hobo)

#convert Date Time to posix 
hobo$DateTime<-format(as.POSIXct(hobo$DateTime,format='%m/%d/%y %I:%M:%S %p'),format="%m/%d/%y %H:%M:%S")
hobo$DateTime<-as.POSIXct(hobo$DateTime,format="%m/%d/%y %H:%M:%S")
```

Read in Odyssey light logger.  
```{r}
light<-read_csv("data/environmental/loggers/calibration/odyssey/15643_20220925_light_calib.csv", skip=7)

colnames(light)<-c("number", "Date", "Time", "Raw.PAR", "Calibrated.PAR")

light <- light %>%
  mutate(DateTime = paste(Date, Time))%>%
  mutate(DateTime = as.POSIXct(DateTime,format="%d/%m/%Y %H:%M:%S"))%>%
  select(DateTime, Calibrated.PAR)%>%
  rename(Light.PAR=Calibrated.PAR)%>%
  mutate(logger="15643")
```

# Merge files together  

First make a list of the SNs for each type of logger before spreading to wide format for each.  
```{r}
hobo_list<-levels(as.factor(hobo$logger))
pendant_list<-levels(as.factor(pendants$logger))
light_list<-levels(as.factor(light$logger))
```

Convert each data frame to wide format.  
```{r}
hobo<-hobo%>%
  spread(value=Temp.C.hobo, key=logger)

#make two for pendants - one for light one for temp
pendants_light<-pendants%>%
  select(!Temp.C.pendant)%>%
  spread(value=Light.lux, key=logger)

pendants_temp<-pendants%>%
  select(!Light.lux)%>%
  spread(value=Temp.C.pendant, key=logger)

light<-light%>%
  spread(value=Light.PAR, key=logger)
```


Something gets messed up here - check tz in all data frames to make sure they are the same upon merging 



```{r}
cal_light<-left_join(pendants_light, light)


cal_temp<-inner_join(hobo, pendants_temp)
```

# Temperature calibrations 

Plot temperature logging over time for each logger and color by black loggers and pendants.  

```{r}
temp_plot<-cal_temp%>%
  gather(value=temp, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=temp, colour=logger))+
  geom_line()+
  theme_classic(); temp_plot

```

Something is off about the time stamps for a few loggers. Look into this. 







The time stamp for loggers 


Calculate mean hobo loogers, this will be our standard temperature.  

Calculate the correlation curve at each observation from the mean hobo loggers for each individual logger.  

Output to a data frame to use in future analysis with these loggers.  


# Light calibrations  

Plot temperature logging over time for each logger and color by odyssey and pendants.  
```{r}
light_plot<-cal_light%>%
  #select(DateTime, `15643`)%>%
  gather(value=light, key=logger, -DateTime)%>%
  
  ggplot(aes(x=DateTime, y=light, colour=logger))+
  geom_line()+
  theme_classic(); light_plot

```

Is the time stamp off for the odyssey logger???? Nee to calculate the offset issue that we are having and re run all scripts from data upload 


Calculate the correlation curve at each observation from the odyssey loggers for each individual logger.  

Output to a data frame to use in future analysis with these loggers.  